### 2015-01-21 sharp-sigma-one-plus.tex
21 ene v0.1

We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
\begin{eqnarray*}
\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
\end{eqnarray*}
where, for every $i\in\{1,\ldots,\hat{k}\}$, 
\begin{eqnarray*}
\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
\end{eqnarray*}
The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously. Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables. Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$. Note that for every formula $\psi(\x)$,
\begin{eqnarray*}
\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
\end{eqnarray*}
The formulas we will be using are $\varphi_i(\x,\X)$, for $i\in\{1,\ldots,k\},$ which are defined as follows: 
\begin{eqnarray*}
\varphi_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\varphi_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
\varphi_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\varphi_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
& \vdots & \\
\varphi_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\varphi_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
\end{eqnarray*}
Note that $k = \hat{k} \cdot g(d).$ Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
\begin{eqnarray*}
\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
\end{eqnarray*}
where $\theta_i(\x) = \theta^{\,i\:(\!\!\!\!\mod g(d))}(\x).$ Also, note that
\begin{eqnarray*}
\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
\end{eqnarray*}

Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on. As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts. Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
\begin{multline*}
\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
\end{multline*}
where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
\begin{multline*}
\psi_i(\X) = \forall\x\:\neg\alpha(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
\end{multline*}
Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
\begin{multline}
\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\varphi_1(\v,\X)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{i-1}(\v,\X)\vee\psi_i(\X)).
\end{multline}
Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$. Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula. As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
\begin{enumerate}
\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$. (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$. (2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$. However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$. And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\exists\x\:\psi_q(\P)$, so necessarily $\A\models\exists\x\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q.$
\end{enumerate}

>>>>

We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
\begin{eqnarray*}
\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
\end{eqnarray*}
where, for every $i\in\{1,\ldots,\hat{k}\}$, 
\begin{eqnarray*}
\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
\end{eqnarray*}
The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously. Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables. Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$. Note that for every formula $\psi(\x)$,
\begin{eqnarray*}
\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
\end{eqnarray*}
We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
\begin{eqnarray*}
\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
& \vdots & \\
\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
& \vdots & \\
\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
\end{eqnarray*}
Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts. We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$

As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$

Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
\begin{eqnarray*}
\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
\end{eqnarray*}
 Also, note that
\begin{eqnarray*}
\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
\end{eqnarray*}

Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on. As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts. Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$. Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula. As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$. (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$. (2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$. However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$. And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$. But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.

Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
